\section{Examples and Definitions}
\label{sec:examples-and-definitions}

There are two ways to view groups, \emph{algebra} and \emph{symmetry}. We first build intuition by looking at an example in symmetry.

\begin{example}[Equilateral Triangle]
    An equilateral triangle has rotational symmetries and reflective symmetries. Note that there is the \emph{trivial symmetry} which is the \emph{identity}, that does not move the triangle and keeps it in place.

    An equilateral triangle has 6 symmetries.
\end{example}

Natural questions arise: what are the symmetries of a square, a regular pentagon, or in general, a regular \(n\)-gon?

We can define 'multiplication' on symmetries, specifically by composing the symmetries:

We note the following features (and non-features) arise in the previous example. \bigbreak

\noindent
\textbf{Important features.}
\begin{itemize}
    \item There is an \emph{identity}.
    \item Every symmetry has an \emph{inverse}.
    \item Symmetries can be \emph{composed} (to form another symmetry).
    \item Composition of symmetries are \emph{associative}.
\end{itemize}

\noindent
\textbf{Non-feature.} The symmetries need not commute.

\subsection{Introduction to the Algebraic Standpoint}
\label{sec:algebraic-standpoint}

We first define what it means to have a binary operation on a set -- this forms the foundation for algebra.

\begin{definition}[Binary Operation]
    \label{def:binary-operation}
    A \emph{binary operation} on a set \(X\) is a function
    \[
        \cdot : X \times X \to X
    \]
    where \(X \times X\) is the \emph{Cartesian product} given by
    \[
        X \times X = \set{\para{x_1, x_2}}{x_1, x_2 \in X}.
    \]
    
    We denote
    \[
        \para{x_1, x_2} \mapsto x_1 \cdot x_2.
    \]
\end{definition}

Now we are ready to define a group as an algebraic structure, using the following axioms.

\begin{definition}[Group]
    \label{def:group}
    A \emph{group} is a triple
    \[
        \para{G, \cdot, e}
    \]
    where
    \begin{itemize}
        \item \(G\) is a set,
        \item \(\cdot\) is a binary operation on \(G\), and
        \item \(e\) is an element of \(G\), denoted as \(e \in G\),
    \end{itemize}
    such that the following 4 axioms are satisfied:
    \begin{enumerate}
        \item \emph{Closure}. For all \(a, b \in G\), \(a \cdot b \in G\).
        \item \emph{Associativity}. For all \(a, b, c \in G\), \(
        \para{\para{a \cdot b} \cdot c} = \para{a \cdot \para{b \cdot c}}\).
        \item \emph{(Right) Identity}. For all \(a \in G\), \(a \cdot e = a\).
        \item \emph{(Right) Inverses}. For all \(a \in G\), there exists some \(b \in G\), such that \(a \cdot b = e\).
    \end{enumerate}
\end{definition}

\begin{remark}[Remark on Closure]
    The closure axiom is redundant, since \(\cdot\) being a binary operation means that \(G\) has to be closed under \(\cdot\)
\end{remark}

A group is an \textbf{algebraic structure with one operation}.

\begin{examples}
    \begin{itemize}
        \item \emph{Equilateral triangle.} The symmetries of an equilateral triangle forms a group, with respect to composition, and the identity symmetry is the identity element.
        \item \emph{Integers.} \(\para{\ZZ, +, 0}\) forms a group.
    \end{itemize}
\end{examples}

Some authors define the identity to be a two-sided identity, and the inverse to be a two-sided inverse. Indeed, this must be the case, and the identity and the inverse of an element must be unique, as in \autoref{prop:inverse-identity}.

\begin{proposition}[Properties on Inverses and Identities]
    \label{prop:inverse-identity}
    Let \(\para{G, \cdot, e}\) be a group. Let \(a, b, b', e' \in G\).

    \begin{enumerate}
        \item \emph{Right-inverses are also left-inverses.} If \(a \cdot b = e\), then \(b \cdot a = e\).
        \item \emph{A right-identity is also a left-identity.} \(e \cdot a = a\).
        \item \emph{Inverses are unique.} If \(a, b, b'\) are such that \(a \cdot b = e\) and \(a \cdot b' = e\), then \(b = b'\).
        \item \emph{The identity is unique.} If \(a \cdot e' = a\), then \(e' = e\).
    \end{enumerate}
\end{proposition}

\begin{proof}
    \begin{enumerate}
        \item We have
        \begin{align*}
            b &= b \cdot e && \text{Identity Axiom} \\
            &= b \cdot \para{a \cdot b} && \text{Assumption} \\
            &= \para{b \cdot a} \cdot b. && \text{Associativity Axiom}
        \end{align*}

        By the Inverse Axiom, there exists \(c \in G\) such that \(b \cdot c = e\).

        We now right-multiply both sides by \(c\), giving
        \begin{align*}
            e &= b \cdot c && \text{Assumption on }c \\
            &= \para{\para{b \cdot a} \cdot b} \cdot c \\
            &= \para{b \cdot a} \cdot \para{b \cdot c} && \text{Associativity Axiom} \\
            &= \para{b \cdot a} \cdot e && \text{Assumption on }c \\
            &= b \cdot a. && \text{Identity Axiom}
        \end{align*}

        \item By the Inverse Axiom and the first part, there exists \(b \in G\) such that \(a \cdot b = b \cdot a = e\).
        
        Now, we have
        \begin{align*}
            e \cdot a &= \para{a \cdot b} \cdot a && \text{Assumption} \\
            &= a \cdot \para{b \cdot a} && \text{Associativity Axiom} \\
            &= a \cdot e && \text{Assumption} \\
            &= a. && \text{Identity Axiom}
        \end{align*}

        \item We have from the previous parts, that
        \begin{align*}
            b' &= e \cdot b' && \text{\autoref{prop:inverse-identity}, Part 2} \\
            &= \para{b \cdot a} \cdot b' && \text{\autoref{prop:inverse-identity}, Part 1, and Assumption} \\
            &= b \cdot \para{a \cdot b'} && \text{Associativity Axiom} \\
            &= b \cdot e && \text{Assumption} \\
            &= b && \text{Identity Axiom}.
        \end{align*}

        \item By the Inverse Axiom, and \autoref{prop:inverse-identity} Part 1, there exists \(b \in G\) such that \(b \cdot a = e\). We are given that \(a \cdot e' = a\), and hence \(b \cdot \para{a \cdot e'} = b \cdot a\).

        Hence,
        \begin{align*}
            e &= \para{b \cdot a} \cdot e' && \text{Associativity Axiom} \\
            &= e \cdot e' \\
            &= e' && \text{\autoref{prop:inverse-identity} Part 2}.
        \end{align*}
    \end{enumerate}    
\end{proof}

Part 3 of \autoref{prop:inverse-identity} tells us that inverses are unique.

\begin{notation}
    If \(a \cdot b = e\), we write \(b = a \inverse\).
\end{notation}

Part 1 of the same proposition tells us that
\[
    a \cdot a \inverse = a \inverse \cdot a = e
\]
which in turn means that
\[
    \para{a \inverse} \inverse = a.
\]

\begin{notation}
    We may extend this notation for
    \begin{itemize}
        \item \(a^0 = e\),
        \item \(a^n = a^{n - 1} \cdot a\) for any \(n \in \NN\).
    \end{itemize}

    Similarly, for negative integers \(n < 0\), we define
    \[
        a^n = \para{a \inverse}^{-n}.
    \]
\end{notation}

We may show that \(a^m \cdot a^n = a^{m + n}\) and \(\para{a^m}^n = a^{mn}\), that is, the exponentiation satisfies the rules we are familiar with.

Recall that it is \emph{not} necessarily true that \(a \cdot b = b \cdot a\) for all \(a, b \in G\).

Hence, it is \emph{not} necessarily true that
\[
    \para{a \cdot b}\inverse = a \inverse \cdot b \inverse.
\]

However, we notice that
\begin{align*}
    \para{a \cdot b} \cdot \para{b \inverse \cdot a \inverse} &= \para{\para{a \cdot b} \cdot b\inverse} \cdot a\inverse \\
    &= \para{a \cdot \para{b \cdot b\inverse}} \cdot a\inverse \\
    &= \para{a \cdot e} \cdot a\inverse \\
    &= a \cdot a\inverse \\
    &= e.
\end{align*}

Hence, we have that
\[
    \para{a \cdot b}\inverse = b \inverse \cdot a \inverse.
\]

We give a special name for those groups which are commutative.

\begin{definition}[Commutativity, Abelian Group]
    \label{def:abelian-group}
    If \((G, \cdot, e)\) is a group, and
    \[
        a \cdot b = b \cdot a
    \]
    for all \(a, b \in G\), then \(G\) is \emph{commutative}, and \(G\) is called an \emph{abelian group}.
\end{definition}

\begin{example}[Trivial Group]
    If \(G = \brce{e}\) and \(\cdot: \para{e, e} \mapsto e\), then \(\para{G, \cdot, e}\) is the \emph{trivial group}.
\end{example}

\begin{examples}[Familiar Examples from Arithemetic]
    \begin{enumerate}
        \item \(\para{\ZZ, +, 0}\), \(\para{\QQ, +, 0}\), \(\para{\RR, +, 0}\) and \(\para{\CC, +, 0}\) are all \emph{abelian} groups.
        
        The inverse of \(x\) in these groups are \(-x\) -- this is known as the \emph{additive notation}.

        \item \(\para{\NN, +, 0}\) is \emph{not} a group, since it does not satisfy the Inverse Axiom.
        
        \item \(\para{\QQ, \times, 1}\) is \emph{not} a group because \(0 \in \QQ\) has no inverse. However, \(\para{\QQ \setminus \brce{0}, \times, 1}\) is an \emph{abelian group}, and we denote this as \(\QQ\multiplicative\).
        
        Similarly, \(\para{\RR \setminus \brce{0}, \times, 1}\) and \(\para{\CC \setminus \brce{0}, \times, 1}\) are as well, and we denote them as \(\RR\multiplicative\) and \(\CC\multiplicative\) respectively.
    \end{enumerate}
\end{examples}

\subsection{Definitions on Finite Groups}
\label{sec:def-finite-groups}

Most examples of groups given above are infinite, apart from the trivial group.

We would like to define the size of a group, and the definition is as in 

\begin{definition}[Order of a Group, Finite Group]
    \label{def:order-group}

    The \emph{order} of a group \(\para{G, \cdot, e}\) is the number of elements of \(G\), denoted by \(\abs{G}\) or \(\size G\).

    If \(\abs{G} < \infinity\), then we say \(\para{G, \cdot, e}\) is a \emph{finite} group.
\end{definition}

\begin{examples}
    \begin{enumerate}
        \item For any \(n \in \NN\), let \(C_n = \set{z \in \CC}{z^n = 1}\), then \(\para{C_n, \times, 1}\) is an abelian group.
        \item Let \(\ZZ_n = \brce{0, 1, \cdots, n - 1}\). We let for \(a, b \in \ZZ_n\), we have
        \[
            a +_n b = (a + b) \modulo n.
        \]

        We can see \(\para{\ZZ_n, +_n, 0}\) is an abelian group.
    \end{enumerate}
\end{examples}

\subsection{Symmetric Groups}
\label{sec:symmetric-groups}

Symmetries on a set (with no additional constraints) are naturally interpreted as permutations, and we give a formal definition below.

\begin{definition}[Bijection, Permutation]
    \label{def:bijection-permutation}
    Let \(X, Y\) be sets. A \emph{bijection} is a map \(X \to Y\) that has an inverse \(g: Y \to X\) such that
    \[
        f \circ g = \identity_Y, g \circ f = \identity_X
    \]
    where
    \[
        \identity_S: S \to S, x \mapsto x
    \]
    for all \(x \in S\) is the \emph{identity map}.

    Here,
    \[
        \para{f \circ g}\para{x} = f \para{g \para{x}}
    \]
    by definition of composition.

    A bijection \(X \to X\) is called a \emph{permutation}.
\end{definition}

\begin{definition}[Symmetry Group]
    \label{def:symmetry-group}
    \(\Sym\para{X}\) is defined to be the set of permutations of a set \(X\).
\end{definition}

To show \(\Sym\para{X}\) is indeed a group, we first show \autoref{lem:composition-associative}.

\begin{lemma}[Composition is Associative]
    \label{lem:composition-associative}

    Consider maps of sets
    \[
        f: W \to X, g: X \to Y, h: Y \to Z,
    \]
    we have
    \[
        \para{h \circ g} \circ f = h \circ \para{g \circ f}.
    \]
\end{lemma}

\begin{proof}
    For any \(w \in W\), we have
    \begin{align*}
        \para{\para{h \circ g} \circ f} \para{w} &= \para{h \circ g} \para{f\para{w}} \\
        &= h \para{g \para{f \para{w}}} \\
        &= h \para{\para{g \circ f} \para{w}} \\
        &= \para{h \circ \para{g \circ f}} \para{w}
    \end{align*}
    as required.
\end{proof}

\begin{proposition}
    \label{prop:symmetry-group}

    For any set \(X\), \(\para{\Sym\para{X}, \circ, \identity_X}\) is a group.
\end{proposition}

\begin{proof}
    Closure is automatic, and associativity follows from \autoref{lem:composition-associative}.

    The identity on \(X\), \(\identity_X\) provides an identity elements.

    Inverses exist by definition.
\end{proof}

\begin{definition}[\(S_n\)]
    \label{def:s-n}

    If \(X = \brce{1, 2, \ldots, n}\), we write \(S_n = \Sym\para{X}\).
\end{definition}

\begin{examples}
    \begin{itemize}
        \item \(S_3\) may represent the groups to rearrange three flower pits.
        \item \(S_{52}\) may represent the group of ways to shuffle a deck of cards.
    \end{itemize}
\end{examples}

\begin{remark}
    We note that \(\abs{S_n} = n!\) by basic combinatorics.
\end{remark}

\begin{notation}
    Writing \(\para{G, \cdot, e}\) for a group is often cumbersome. Henceforth, we will normally just write \(G\) for the group.
    
    If we would like to emphasise that \(e\) is the identity element of \(G\), we will write \(e_G\) for \(e\).

    Likewise, we can write \(\cdot_G\) for the operation on \(G\).
\end{notation}

\subsection{Subgroups}
\label{sec:subgroups}

Sometimes we want to restrict our attention to smaller groups rather than the whole group, for instance, \(\ZZ\) inside \(\RR\), or the rotations that preserve a triangle, instead of all symmetries of the plane.

\begin{definition}[Subgroup]
    \label{def:subgroup}

    Let \(G\) be a group, and \(H \subseteq G\). If we have
    \begin{itemize}
        \item \emph{Identity.} \(e \in H\);
        \item \emph{Closure.} \(a \cdot b \in H\) for all \(a, b \in H\);
        \item \emph{Inverse.} \(a \in H\) implies \(a\inverse \in H\);
    \end{itemize}
    we say \(H\) is a \emph{subgroup} of \(G\), and we write \(H \leq G\).
\end{definition}

\begin{remark}
    If \(G\) is a group, then any subgroup \(H \leq G\) is itself a group. Note associativity is not required -- it is inherited from \(G\).
\end{remark}

\begin{examples}
    \begin{enumerate}
        \item Every group \(G\) is a subgroup of itself.

        \item For any group \(G\),
        \[
            1_G = \brce{e_G}
        \]
        is the \emph{trivial subgroup}.

        \item \(\ZZ \subseteq \QQ \subseteq \RR \subseteq \CC\), and in fact
        \[
            \ZZ \leq \QQ \leq \RR \leq \ZZ.
        \]

        \item For \(n = 0, 1, 2, 3, \cdots,\) let
        \[
            n\ZZ = \set{nk}{k \in \ZZ} = \brce{\cdots, -2n, -n, 0, n, 2, \cdots}.
        \]

        Now, we have
        \begin{itemize}
            \item \(0 \in \ZZ\).
            \item Let \(x = nk, y = nl \in n\ZZ\) for \(k, l \in \ZZ\).
            
            Then \(x + y = nk + nl = n \para{k + l} \in n\ZZ\).
            \item Let \(x = nk \in n\ZZ\) for \(k \in \ZZ\).
            
            Then \(-x = -nk = n \cdot \para{-k} \in n\ZZ\).
        \end{itemize}

        Hence, \(n\ZZ \leq \ZZ\).

        We will see in \autoref{prop:subgroups-of-integers} that all subgroups of \(\ZZ\) are of this form.
    \end{enumerate}
\end{examples}

\begin{remark}
    All subgroups of a group \(G\) that are not trivial (\(1_G\)) or the group \(G\) itself are referred to as its \emph{proper subgroups}.
\end{remark}

\begin{proposition}
    \label{prop:subgroups-of-integers}

    If \(H \leq \ZZ\), then \(H = n\ZZ\) for some \(n = 0, 1, 2, \cdots\).
\end{proposition}

\begin{proof}
    If \(H = \brce{0}\) then \(H = 0\ZZ\) is indeed of this form.

    Otherwise, \(H \neq \brce{0}\), we may choose \(n \in H \setminus \brce{0}\) to be the smallest positive \(n \in H\).

    Note that if \(x \in H\) and \(x < 0\), then \(-x \in H\) and \(x > 0\). Therefore, such positive number \(n\) must exist.

    By induction, we may show that \(nk \in H\) for all \(k = 1, 2, 3, \cdots\), and \(-n \in H\), so \(-nk \in H\) for all \(k = 1, 2, 3, \cdots\).

    This gives us \(n\ZZ \subseteq H\), and we next show \(n\ZZ = H\) by contradiction. Suppose \(n \ZZ \neq H\), so there is some \(x \in H\) such that \(x \notin n\ZZ\).

    Dividing by \(n\) and taking the remainder (integer division), we get \(x = nq + r\) for some integers \(q, r \in \ZZ\), where \(0 < r < n\).

    But we know \(r = x - nq\) and \(x \in H, nq \in n\ZZ \subseteq H\) so \(r \in H\), and this contradicts that \(n\) is the smallest.

    So \(n\ZZ = H\) as desired.
\end{proof}