\section{Finite Groups}
\label{sec:finite-groups}

We have already seen several useful theorems on finite groups, e.g. \autoref{thm:lagrange} (Lagrange's Theorem), \autoref{cor:orbit-stabiliser} (Orbit--Stabiliser Theorem), \autoref{thm:cayley} (Cayley's Theorem) and \autoref{thm:cauchy} (Cauchy's Theorem).

In this section we will be thinking of small examples of finite groups. Namely, we proceed na\"{i}vely, classifying all groups up to order \(8\).

\subsection{Groups of Order 1, 2, 3, 5 and 7}
\label{sec:groups-of-order-1-2-3-5-7}

\begin{itemize}
    \item If \(\abs{G} = 1\), then \(G\) must be trivial: \(G \isomorphic \trivialgp\).
    \item If \(\abs{G} = 2, 3, 5, 7\), since those are all primes, by \autoref{cor:group-prime-order} (groups of prime order are cyclic), \(G \isomorphic C_{\abs{G}}\).
\end{itemize}

\subsection{Groups of Order 4}
\label{sec:groups-of-order-4}

We have \(C_4\) as an example of a group of order \(4\). For another group of this order, we first define the following.

\begin{definition}[Direct Product]
    \label{def:direct-product}%
    If \(G, H\) are groups, the \emph{direct product} is
    \[
        G \times H = \set{\para{g, h}}{g \in G, h \in H}
    \]
    equipped with the group operation
    \[
        \para{g_1, h_1} \cdot_{G \times H} \para{g_2, h_2} = \para{g_1 \cdot_G g_2, h_1 \cdot_H h_2}.
    \]

    Note that the element \(e_G, e_H\) is the identity, and for the inverse
    \[
        \para{g, h}\inverse = \para{g\inverse, h\inverse}.
    \]
\end{definition}

The direct product of two groups is a group, and it is easy to see for any group \(G, H, K\),
\begin{itemize}
    \item \(G \times H \isomorphic H \times G\): the direct product is \emph{commutative} up to isomorphism;
    \item \(\para{G \times H} \times K \isomorphic G \times \para{H \times K}\): the direct product is \emph{associative} up to isomorphism;
    \item \(\trivialgp \times G \isomorphic G \times \trivialgp \isomorphic G\): the trivial group is \textbf{an} \emph{identity} up to isomorphism (for infinite groups, \(G \times H \isomorphic G\) \textbf{does not require} \(H\) to be trivial, for example \(\RR \times \RR \isomorphic \RR\), since they are isomorphic when viewed as \(\QQ\)-vector spaces);
    \item If \(G \isomorphic H\), then \(G \times K \isomorphic H \times K\): we may replace elements in the direct product up to isomorphism;
    \item \(\abs{G \times H} = \abs{G} \times \abs{H}\): the order of the direct product is the product of the orders;
    \item \(\ord\para{g, h} = \lcm\brce{\ord\para{g}, \ord\para{h}}\): the order of an element is the least common multiple of the order of the separate elements.
\end{itemize}

\begin{remark}
    The direct product \(H_1 \times H_2\) provides a \emph{projection map}, defined as
    \[
        \function{\pi}{H_1 \times H_2}{H_1}{\para{h_1, h_2}}{h_1}
    \]
    and obviously this map is surjective with kernel
    \[
        \ker \pi = \trivialgp_{H_1} \times H_2 \isomorphic H_2.
    \]
\end{remark}

\begin{example*}[The Klein--4 Group]
    The group
    \[
        K_4 = C_2 \times C_2
    \]
    is called the \emph{Klein--4 Group}. It can be thought of the dihedral group \(D_4\).

    For any \(\para{a, b} \in K_4\), we have
    \[
        \para{a, b}^2 = \para{a^2, b^2} = \para{e, e} = e,
    \]
    so every element has order \(1\) or order \(2\). In particular, \(K_4 \notisomorphic C_4\).
\end{example*}

To show these are the only groups of order \(4\), it is useful to introduce a theorem regarding direct products.

\begin{theorem}[Direct Product Theorem]
    \label{thm:dpt-direct-product-theorem}%
    If \(H_1, H_2 \subgroup G\), and
    \begin{enumerate}
        \item \(H_1 \cap H_2 = \brce{e}\): \(H_1\) and \(H_2\) \emph{intersect trivially};
        \item \(h_1 h_2 = h_2 h_1\) for all \(h_1 \in H_1, h_2 \in H_2\): elements in \(H_1\) and \(H_2\) \emph{commute}; and
        \item \(G = H_1 H_2\): for all \(g \in G\), \(g = h_1 h_2\) for some \(h_1 \in H_1\) and \(h_2 \in H_2\);
    \end{enumerate}
    then \(G \isomorphic H_1 \times H_2\).
\end{theorem}

\begin{proof}
    We consider the na\"{i}ve map
    \[
        \function{\phi}{H_1 \times H_2}{G}{\para{h_1, h_2}}{h_1 h_2.}
    \]

    We show that \(\phi\) is an isomorphism.

    \emph{\(\phi\) is a homomorphism.} For any \(\para{h_1, h_2}, \para{h_1', h_2'} \in H_1 \times H_2\), we have
    \begin{align*}
        \phi\para{\para{h_1, h_2}} \phi\para{\para{h_1', h_2'}} &= \para{h_1 h_2} \para{h_1' h_2'} \\
        &= \para{h_1 h_1'} \para{h_2 h_2'} && \text{by condition (iii)} \\
        &= \phi\para{\para{h_1 h_1', h_2 h_2'}} \\
        &= \phi\para{\para{h_1, h_2} \cdot \para{h_1', h_2'}}.
    \end{align*}

    \emph{\(\phi\) is injective.} Recall this is equivalent to showing that the kernel \(\ker \phi = \trivialgp\) is trivial.

    Suppose \(\para{h_1, h_2} \in H_1 \times H_2\) is such that \(\phi\para{\para{h_1, h_2}} = e = h_1 h_2\), i.e. \(\para{h_1, h_2} \in \ker\phi\).

    Then we have
    \[
        H_1 \ni h_1 = h_2\inverse \in H_2
    \]
    so \(h_2\inverse = h_1 \in H_1 \cap H_2 = \trivialgp\) by condition (i). Hence, \(h_1 = h_2 = e\) and \(\para{h_1, h_2} = \para{e, e}\), showing \(\ker \phi = \trivialgp\) as desired.

    \emph{\(\phi\) is surjective.} This is from condition (iii).

    This shows \(\phi\) is an isomorphism, as desired.
\end{proof}

\begin{remark}
    Condition (iii) is to show \(\phi\) is surjective, and in the case where \(H_1, H_2 \leq G\) is finite, we have if \(H_1 \cap H_2 = \trivialgp\), then
    \[
        \abs{\set{h_1 h_2}{h_1 \in H_1, h_2 \in H_2}} = \abs{H_1 H_2} = \abs{H_1} \abs{H_2},
    \]
    so we may replace condition (iii) with \(\abs{G} = \abs{H_1} \abs{H_2}\) to show that the mapping is surjective.
\end{remark}

\begin{remark}
    For any group \(G \isomorphic H_1 \times H_2\), we may choose the natural copies of \(H_1\) and \(H_2\) in \(G\), \(H_1'\) and \(H_2'\):
    \[
        H_1' = H_1 \times \trivialgp = \set{\para{h_1, e}}{h_1 \in H_1} \isomorphic H_1
    \]
    and
    \[
        H_2' = \trivialgp \times H_2 = \set{\para{e, h_2}}{h_2 \in H_2} \isomorphic H_2,
    \]
    and we have \(G \isomorphic H_1' \times H_2'\). \(H_1'\) and \(H_2'\) here satisfy the condition for \autoref{thm:dpt-direct-product-theorem} (Direct Product Theorem). Therefore, this theorem is, in a sense, `if and only if', up to isomorphism.
\end{remark}

Now we are ready to classify all groups of order 4.

\begin{lemma}[Groups of Order 4]
    \label{lem:groups-of-order-4}%
    If \(\abs{G} = 4\), then \(G \isomorphic C_4\) or \(G \isomorphic C_2 \times C_2\).
\end{lemma}

\begin{proof}
    By \autoref{cor:order-of-element-divides} (Corollary of Lagrange's Theorem), every non-trivial element in \(G\) has order \(2\) or \(4\).
    
    \begin{itemize}
        \item If there is some \(g \in G\) such that \(\ord\para{g} = 4\), then \(G = \ang{g} \isomorphic C_4\).
        \item Otherwise, every non-trivial element has order \(2\). Let \(a, b \in G\) be distinct elements such that \(\ord{a} = \ord{b} = 2\).
        
        Let \(H_1 = \ang{a} \isomorphic C_2\), \(H_2 = \ang{b} \isomorphic C_2\). We aim to show that \(G \isomorphic H_1 \times H_2 \isomorphic C_2 \times C_2\).

        First, \(H_1 \cap H_2 = \trivialgp_{G}\) since \(a \neq b\), giving condition (i). Then, \(\abs{G} = \abs{H_1} \abs{H_2}\), which gives condition (iii).

        Finally, since \(\ord\para{ab} = 2\) (since \(ab \neq e\)), then
        \[
            ab = \para{ab}\inverse = b\inverse a\inverse = ba,
        \]
        as required for (ii).

        Therefore, \(G \isomorphic C_2 \times C_2\) by \autoref{thm:dpt-direct-product-theorem} (Direct Product Theorem), as desired.
    \end{itemize}
\end{proof}

It is useful to see when/whether a direct product of cyclic groups is cyclic -- in the example above \(C_2 \times C_2 \notisomorphic C_4\), but we will soon see that \(C_2 \times C_3 \isomorphic C_6\).

\begin{theorem}[Chinese Remainder Theorem]
    \label{thm:crt-chinese-remainder-theorem}%
    If \(\gcd\brce{m, n} = 1\), then
    \[
        C_m \times C_n \isomorphic C_{mn}.
    \]
\end{theorem}

\begin{proof}
    Let \(C_{mn} = \ang{g}\). Let
    \[
        H_1 = \ang{g^n} \isomorphic C_m, H_2 = \ang{g^m} \isomorphic C_n,
    \]
    we check the hypothesis of \autoref{thm:dpt-direct-product-theorem} (Direct Product Theorem).

    \emph{(i).} Note that \(g^k \in H_1\) if and only if \(n \divides k\), and similarly, \(g^k \in H_2\) if and only if \(m \divides k\).

    Hence, \(g^k \in H_1 \cap H_2\) if and only if \(m \divides k\) and \(n \divides k\). This is equivalent to
    \[
        \lcm\brce{m, n} \divides k,
    \]
    and since \(\gcd\brce{m, n} = 1\), we have \(\lcm\brce{m, n} = mn\), and hence this is equivalent to
    \[
        mn \divides k,
    \]
    which is equivalent to \(g^k = e\).

    \emph{(ii).} Cyclic groups are abelian.

    \emph{(iii).} \(\abs{C_{mn}} = \abs{C_m} \abs{C_n}\) so this is immediate.
\end{proof}

\subsection{Groups of Order 6}
\label{sec:groups-of-order-6}

We may guess the groups of order \(6\) are the dihedral group \(D_6\), and the cyclic group \(C_6 \isomorphic C_2 \times C_3\). In fact, through some experiment, we may see that they are the only ones.

\begin{lemma}[Groups of Order 6]
    \label{lem:groups-of-order-6}%
    If \(\abs{G} = 6\), then \(G \isomorphic C_6\) or \(G \isomorphic D_6\).
\end{lemma}

\begin{proof}
    By \autoref{thm:cauchy} (Cauchy's Theorem), there are \(r, s \in G\) such that \(\ord\para{r} = 3, \ord\para{s} = 2\).

    Since \(\abs{\ang{r}} = 3\), we have from \autoref{thm:lagrange} (Lagrange's Theorem), that \(\abs{G : \ang{r}} = 2\), i.e. there are two cosets of \(\ang{r}\). One of them is \(\ang{r}\), and since \(s \in \ang{r}\), the other left coset is \(s\ang{r}\). Similarly, for right cosets, the right coset other than the trivial one is \(\ang{r}s\). Therefore, we have
    \[
        s\ang{r} = G \setminus \ang{r} = \ang{r} s.
    \]

    Hence, \(sr = r^i s\) for some \(i = 0, 1, 2\).

    \begin{itemize}
        \item \(i = 0\). Then \(sr = s\), so \(r = e\) which is impossible since \(\ord\para{r} = 3\).
        \item \(i = 1\). Then \(sr = rs\), so elements in \(\ang{r}\) and \(\ang{s}\) commute. By \autoref{thm:dpt-direct-product-theorem} (Direct Product Theorem),
        \[
            G \isomorphic \ang{r} \times \ang{s} \isomorphic C_3 \times C_2 \isomorphic C_6.
        \]
        \item \(i = 2\). Then \(sr = r^2 s = r\inverse s\) give the \emph{dihedral relation}. Therefore, by \autoref{prop:group-isomorphic-dihedral}, we have \(G \isomorphic D_6\).
    \end{itemize}
\end{proof}

\begin{remark}
    Another obvious group of order \(6\) is \(S_3\), but in fact \(S_3 \isomorphic D_6\) since it is non-abelian with order \(6\). (Consider permutations on vertices of triangle.)
\end{remark}

\subsection{Groups of Order 8}
\label{sec:groups-of-order-8}

We already know the groups \(C_2 \times C_2 \times C_2\), \(C_2 \times C_4\) and \(C_8\) as abelian groups of order 8, and \(D_8\) which is non-abelian, and by examining the order of elements, we can easily see none of them are isomorphic to each other.

Another example of a group of order 8 is the quaternion group \(Q_8\).

\begin{examples*}[The Quaternion Group \(Q_8\)]
    The \emph{quaternion group} is
    \begin{align*}
        Q_8 &= \brce{
            \begin{pmatrix}
                \pm 1 & 0 \\ 0 & \pm 1
            \end{pmatrix},
            \begin{pmatrix}
                \pm i & 0 \\ 0 & \mp i
            \end{pmatrix},
            \begin{pmatrix}
                0 & \pm 1 \\ \mp 1 & 0
            \end{pmatrix},
            \begin{pmatrix}
                0 & \pm i \\ \pm i & 0
            \end{pmatrix}} \\
        &\leq \SU\para{2} \leq \SL_2\para{\CC} \leq \GL_2\para{C}.
    \end{align*}

    This is a group under matrix multiplication.

    We usually use Hamilton's notation to simplify this: let
    \begin{align*}
        1 &= \begin{pmatrix}
            1 & 0 \\ 0 & 1
        \end{pmatrix}, \\
        i &= \begin{pmatrix}
            i & 0 \\ 0 & -i
        \end{pmatrix},\\
        j &= \begin{pmatrix}
            0 & 1 \\ -1 & 0
        \end{pmatrix},\\
        k &= \begin{pmatrix}
            0 & i \\ i & 0
        \end{pmatrix},
    \end{align*}
    then we have
    \[
        Q_i = \brce{\pm 1, \pm i, \pm j, \pm k}
    \]
    satisfying
    \begin{itemize}
        \item \(i^2 = j^2 = k^2 = -1\),
        \item \(\para{-1} i = -i\), etc.,
        \item \(ij = k, jk = i, ki = j\), and
        \item \(\para{-1}\) commutes with everything.
    \end{itemize}
\end{examples*}

Note since \(ji = -k \neq k = ij\), \(Q_8\) is \emph{non-abelian}, so
\[
    Q_8 \notisomorphic C_2 \times C_2 \times C_2, C_2 \times C_4, C_8,
\]
and since \(D_8\) has 5 elements of order 2, while \(Q_8\) only has 1 (\(-1\)), we see
\[
    Q_8 \notisomorphic D_8.
\]

In fact these are all groups of order 8.

\begin{lemma}[Groups of Order 8]
    \label{lem:groups-of-order-8}%
    If \(\abs{G} = 8\), then \(G\) is isomorphic to one of the following:
    \[
        C_2 \times C_2 \times C_2, C_2 \times C_4, C_8, D_8, Q_8.
    \]
\end{lemma}

\begin{proof}
    By \autoref{cor:order-of-element-divides} (Corollary of Lagrange's Theorem), every element of \(G\) has order \(1\), \(2\), \(4\) or \(8\). We conduct casework on the order of the elements in the group.

    \begin{itemize}
        \item If there is an element of order \(8\), then \(G \isomorphic C_8\).
        
        \item Suppose that every non-trivial element has order \(2\). Then choosing non-trivial elements
        \[
            a, b \neq a, c \neq a, c \neq b, c \neq ab
        \]
        gives by \autoref{thm:dpt-direct-product-theorem} (Direct Product Theorem) that
        \begin{align*}
            G &\isomorphic \ang{a} \times \ang{b} \times \ang{c} \\
            &\isomorphic C_2 \times C_2 \times C_2.
        \end{align*}

        \item We now suppose that there is an \(a \in G\) such that \(\ord\para{a} = 4\). Let \(b \in G \setminus \ang{a}\), and we have from \autoref{thm:lagrange} (Lagrange's Theorem) that
        \[
            \abs{G : \ang{a}} = 2
        \]
        so
        \[
            b\ang{a} = G \setminus \ang{a} = \ang{a} b.
        \]

        Therefore,
        \[
            b a = a^i b
        \]
        for some \(i = 0, 1, 2, 3\).

        \begin{itemize}
            \item If \(i = 0\), we have \(ba = eb = e\), which gives \(a = e\). This is impossible.
            \item If \(i = 1\), we have \(ba = ab\), and hence by induction, \(b a^j = a^j b\) for all \(j\), therefore \(G\) is abelian. We leave this case here.
            \item If \(i = 2\), we have \(ba = a^2 b\), and hence
            \[
                b a b\inverse = a^2,
            \]
            and the left-hand side has order 4:
            \[
                \ord\para{b a b \inverse} = \ord\para{a} = 4,
            \]
            while the right-hand side has order 2. This is impossible.
            \item If \(i = 3\), this gives \(ba = a^3 b = a\inverse b\), the dihedral relation. We leave this case here.
        \end{itemize}

        On the other hand, note that if \(b^2 \in b \ang{a}\), then \(b \in \ang{a}\), which is not true. So \(b^2 \in \ang{a}\), and we get several cases here.
        \begin{itemize}
            \item \(b^2 = e\). We leave this for future discussion.
            \item \(b^2 = a\). Hence, \(\ord\para{b} = 8\), and hence \(G = \ang{b} \isomorphic C_8\).
            \item \(b^2 = a^2\). We leave this for future discussion.
            \item \(b^2 = a^3 = a\inverse\). Hence, \(\ord\para{b} = 8\), and hence \(G = \ang{b} \isomorphic C_8\).
        \end{itemize}

        Now, we have 4 remaining combinations to discuss.
        \begin{itemize}
            \item \(G\) being abelian, and \(b^2 = e\). By \autoref{thm:dpt-direct-product-theorem} (Direct Product Theorem), since \(\ang{b} \isomorphic C_2\), \(\ang{a} \isomorphic C_4\), and \(\ang{a} \cap \ang{b} = \brce{e}\), we have
            \[
                G \isomorphic C_2 \times C_4.
            \]
            \item \(G\) being abelian, and \(b^2 = a^2\), and therefore \(\para{a b \inverse}^2 = e\), so \(\ord \para{a b \inverse} = 2\). Similarly, by \autoref{thm:dpt-direct-product-theorem} (Direct Product Theorem), we have \(G \isomorphic C_2 \times C_4\).
            \item \(\ord\para{a} = 4\), \(\ord\para{b} = 2\), \(ba = a\inverse b\). By \autoref{prop:group-isomorphic-dihedral}, we see \(G \isomorphic D_8\).
            \item \(\ord\para{a} = 4\), \(b^2 = a^2\) and \(b a = a \inverse b\). We set \(i = a\), \(j = b\), \(k = ab\), \(-1 = a^2 b^2\), then this defines an isomorphism \(Q_8 \to G\) as in \autoref{tab:quartonian-group-isomorphism}.

            \begin{table}[ht!]
                \centering
                \begin{tabular}{cc}
                    \(Q_8\) & \(G\) \\
                    \hline
                    \(e\) & \(e\) \\
                    \(i\) & \(a\) \\
                    \(-1\) & \(a^2\) \\
                    \(-i\) & \(a^3\) \\
                    \(j\) & \(b\) \\
                    \(k\) & \(a b\) \\
                    \(-j\) & \(a^2 b\) \\
                    \(-k\) & \(a^3 b\)
                \end{tabular}
                \caption{Isomorphism of Elements in \(Q_8\) and \(G\)}
                \label{tab:quartonian-group-isomorphism}
            \end{table}
        \end{itemize}
    \end{itemize}
\end{proof}

In summary, \autoref{tab:groups-up-to-order-8} gives a table of all groups up to order 8, up to isomorphism.

\begin{table}[ht!]
    \centering
    \begin{tabular}{cl}
        \(\abs{G}\) & \(G\) \\
        \hline
        1 & \(\trivialgp\) \\
        2 & \(C_2 \isomorphic S_2\) \\
        3 & \(C_3 \isomorphic A_3\) \\
        4 & \(C_4, C_2 \times C_2 \isomorphic D_4 \isomorphic K_4\) \\
        5 & \(C_5\) \\
        6 & \(C_6 \isomorphic C_2 \times C_3, D_6 \isomorphic S_3\) \\
        7 & \(C_7\) \\
        8 & \(C_8, C_4 \times C_2, C_2 \times C_2 \times C_2, D_8, Q_8\)
    \end{tabular}
    \caption{Groups of Order up to 8 up to Isomorphism}
    \label{tab:groups-up-to-order-8}
\end{table}